coDoc: A Tool for Managing Relationships between Code and Documentation

### 2013-02-22 P4-5-page-paper/paper.tex
backup

\title{coDoc: A Tool for Managing Relationships between Codes and Documents}
\author{Dejun Qian\\electronseu@gmail.com}

@@

\begin{abstract}
When writing and testing software, 
we need the documents in hand.
This is extremely true for software that is document intensive, 
such as hardware driver and virtual device modules.
This paper presents coDoc, 
an integrated development environment for managing relationships between codes and documents, as well as developing the codes. 
We first describe the architecture and the features of coDoc.
We then discuss several design decisions that we faced in developing coDoc and the trade-offs in making these decisions.
To provide perspective,
we report our experiences in mapping two virtual network interface card (NIC) modules to the real NIC documents.
\end{abstract}

% Include the motivation for your work, your approach, your goals, and a short summary of what you found/achieved/tested.
\section{Introduction}
\label{sec:introduction}
Documents play an important role in software engineering process,
from requirement analysis to architecture design,
from code writting to software testing,
from software maintainence to software upgrading.
Software that relates closely to hardware devices is extremely document intensive.
Examples of these software could be bootloader, 
Operating System (OS) Hardware Abstraction Layer (HAL),
virtual devices of virtual machines.
These software should conform to hardware documents to operate correctly.
A typical embedded CPU usually have documents with thousands of pages.
A computer system usually consistes of over ten hardware modules.
To effectively maintain the relationships between the codes and the documents is really important for developing and verification of these software.

Much work has been done to deal with the relationships between Application Programming Interface (API) documents and their client codes \cite{Pandita_inferring_2012}.
API documents are usually well indexed by the functions or the classes.
When we are interested in a piece of code in the client software,
we can easily find the related API documents by the name of the functions that the piece of code calls.
The relationships between the client codes and the API documents can be easily constructed by using the function names as the keywords.
However, the codes related to hardware and their documents are usually related in different ways.
The hardware documents usually talk about how to use the hardware,
more specifically how to program the registers inside the hardware modules.
The hardware documents differ from the API documents in the following ways,
\begin{itemize}
\item the hardware documents are usually published as pdf files, while the API documents are usually html files.
\item the hardware documents are composed of texts, tables and diagrams, while the API documents are usually texts only.
\item the hardware docuemnts are not orgnized by functions or classes, which are basic elements of codes.
\item hard for the computer to relate the hardware document pieces to the code pieces automatically.
\end{itemize}

These differences make the constructing of the relationships between the hardware related codes and their documents really challenging.
To our knowlege, there is no tools exist to address this problem.
A simple workaround used is to include the positions of the related document pieces in the documents as comments the codes.
However, this method can only handle simple situations.
If there are thousands of these comments cluttered in the codes,
it will be very hard to manage them.
This paper presents coDoc, a tool to easily create, maintain, and display the relationships.
% The initial version\footnote{\texttt{\url{git@github.com:derekqian/coDoc.git}}} has been released.
The key features of coDoc are as follows,
\begin{itemize}
\item provide the ability to display codes and documents.
\item provide the ability to develop codes.
\item select code pieces based on syntax tree of the code.
\item code selections are stable to the changes of unrelated codes.
\item create relationships between codes and documents.
\item manage and display the relationships.
\end{itemize}

Using coDoc, we have created relationships for two virtual NIC devices in QEMU.
These two devices are DIO and E100.

>>>>

\title{coDoc: A Tool for Managing Relationships between Code and Documentation}
\author{Dejun Qian\\electronseu@gmail.com}

@@

\begin{abstract}
When writing and testing software, 
lacking of the ability to efficiently refer to the related documentation hinders the productivity largely.
This is particularly true for software that is documentation-intensive, 
such as hardware drivers and virtual device modules.
This paper presents coDoc, 
an integrated development environment for managing relationships between code and documentation, while developing the code. 
We describe the architecture and features of coDoc,
and discuss several design decisions that we faced in developing coDoc and the trade-offs involved in making these decisions.
We report our experiences in mapping two virtual network interface card (NIC) modules to the real NIC documents by coDoc.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
Documents play an important role in the software engineering process,
from requirements analysis to architecture design,
from coding to software testing,
from software debugging to software maintenance.
Software that relates closely to hardware devices is extremely document intensive. % explain why
Examples of such software include bootloaders, 
Operating System (OS) Hardware Abstraction Layers (HAL),
and virtual devices for virtual machines.
Such software must follow the hardware documentation to operate correctly.
A typical embedded CPU might have documentation consisting of thousands of pages.
A computer system usually consists of more than ten hardware modules.
Effectively maintaining the relationships between the code and the documentation is important for developing and verifying such software. % which software

Much work has been done to handle the relationships between Application Programming Interface (API) documents and their client code \cite{Pandita_inferring_2012} \cite{wei_inferring_2011}.
API documents are usually well indexed by the functions or the classes that they define.
When we are interested in a piece of code in the client software,
we can easily find the related API documents by the names of the functions that the piece of code calls.
% The relationships between the client codes and the API documents can be easily constructed by using the function names as the keywords.
However, the code relating to hardware and its documentation usually relate in different ways. % how?
The hardware documents usually talk about how to use the hardware,
more specifically how to program the registers inside the hardware modules.
The hardware documents differ from the API documents in the following ways.
\begin{itemize}
\item the hardware documents are usually published as pdf files, while the API documents are usually html files. % matters?
\item the hardware documents are composed of text, tables and diagrams, while the API documents are usually text only.
\item the hardware docuemnts are not orgnized by functions or classes, which are the basic elements of code, but instead by hardware modules.
\item as a consequence of these differences, it is hard for the computer to relate the hardware document pieces to the code pieces automatically.
\end{itemize}

% These differences make the constructing of the relationships between the hardware related codes and their documents really challenging.
To our knowlege, there are no tools to address this problem.
A simple workaround is to include references to positions of the related documentation as comments in the codes.
However, this can only handle simple situations.
If there are thousands of such comments cluttered in the code,
it will be very hard to manage them.
This paper presents coDoc, a tool to create, maintain, and display the relationships between code and documentation.
% The initial version\footnote{\texttt{\url{git@github.com:derekqian/coDoc.git}}} has been released.
The key features of coDoc are as follows:
\begin{itemize}
\item the ability to display both code and documents;
\item the ability to write and edit code;
\item code pieces selection based on syntax tree of the code;
\item stable code selections regarding to the changes of unrelated codes;
\item creation of relationships between codes and documents;
\item relationships management and display.
\end{itemize}

Using coDoc, we have created relationships between two virtual NIC devices in QEMU and the related documents.
These two devices are DIO \cite{dio} and E100.
